---
title: "Exercise 1: Titanic"
output: pdf_document
---
# Exercise 1: Titanic 
```{r, echo=FALSE}
# nolint start: indentation_linter.
data <- read.table("../data/titanic.txt", header = TRUE)
data <- data.frame(data)
data$PClass <- factor(data$PClass)
data$Sex <- factor(data$Sex)

data_not_na <- data[!is.na(data$Age), ]
data_na <- data[is.na(data$Age), ]
age <- data_na$Age
p_class <- data$PClass
sex <- data$Sex
survived <- data$Survived
```

## Part A
```{r, echo=FALSE}
colors <- c("green", "orange", "brown")
plot_labels <- c("Age defined", "Class", "Sex")

par(mfrow = c(1, 4))
par(mar = c(2, 2, 4, 2), cex = 0.75)
barplot(
    matrix(c(
        sum(survived == 1), sum(survived == 0)
    ), ncol = 1),
    names.arg = "Survived",
    col = colors,
    beside = TRUE,
    ylim = c(0, 1000)
)
legend(x = "topleft", fill = colors, legend = c("Survived", "Died"))
barplot(
    matrix(c(
        sum(sex == "male"), sum(sex == "female")
    ), ncol = 1),
    names.arg = "Sex",
    col = colors,
    beside = TRUE,
    ylim = c(0, 1000)
)
legend(x = "topleft", fill = colors, legend = c("Male", "Female"))
barplot(
    matrix(c(length(data_not_na$Age), length(data_na$Age)), ncol = 1),
    names.arg = "Age",
    col = colors,
    beside = TRUE,
    ylim = c(0, 1000)
)
legend(x = "topleft", fill = colors, legend = c("Available", "Missing"))
barplot(
    matrix(c(
        sum(p_class == "1st"), sum(p_class == "2nd"),
        sum(p_class == "3rd")
    ), ncol = 1),
    names.arg = "Class",
    col = colors,
    beside = TRUE,
    ylim = c(0, 1000)
)
legend(x = "topleft", fill = colors, legend = c("1st", "2nd", "3rd"))
mtext("Titanic data summary", side = 3, line = -2, outer = TRUE)
```

```{R, fig.asp = .3, echo=FALSE}
par(mfrow = c(1, 1))
par(mar = c(2, 1, 1.5, 1))
boxplot(matrix(c(data_not_na$Age)), horizontal = TRUE, hight = 1)
mtext("Known age distribution", side = 3, line = -1, outer = TRUE)
```
The data for this exercise is a part of the passenger manifest for the Titanic. The dataset covers the names, passenger class, survival status and, when available, the age of 1313 passengers. A summary of the data is given above. Because the age of some passengers is unavailable, we will use a which does not include these passengers when their age is required. 

```{R}
age_glm <- glm(Survived ~ Age, data = data_not_na, family=binomial)
summary(age_glm)
sex_glm <- glm(Survived ~ Sex, data = data, family=binomial)
summary(sex_glm)
exp(0.69315-2.30118)
cls_glm <- glm(Survived ~ PClass, data = data, family=binomial)
summary(cls_glm)
exp(0.4029-0.7052)
exp(0.4029-1.8265)
```

```{r, echo=F}
# nolint end
```