---
title: "Assignment 1"
author: "Paul Jungnickel, Wessel Beumer, Kushnava Singha, group 10"
date: "23 02 2025"
output: pdf_document
highlight: tango
---
`r options(digits=3)`

## Exercise 1
**a)** 
```{r, fig.height=3, fig.width=6}

cholesterol <- read.table("../data/cholesterol.txt", header = TRUE);
chol_frame <- data.frame(cholesterol);

```


```{r, fig.height=3, fig.width=6}

par(mfrow=c(1,2)); 
p1 <- hist(chol_frame$Before, 8);
p2 <- hist(chol_frame$After8weeks, 6)
qqnorm(chol_frame$Before);
qqline(chol_frame$Before);
qqnorm(chol_frame$After8weeks);
qqline(chol_frame$After8weeks); # two QQ-plots next to each other
```

The histograms of the two samples both appear normal, even though the sample size is very small. The Q-Q plots confirm this.




`cor(chol_frame$Before, chol_frame$After8weeks)` `r cor(chol_frame$Before, chol_frame$After8weeks)`

The correlation between the two samples is almost one i.e. a high cholesterol at the start also implies a high cholesterol level after 8 weeks.



**b)** 
The experiment outcomes are paired , since the same person is studied at two time units


```{r, fig.height=3, fig.width=6}


t.test(chol_frame$Before, chol_frame$After8weeks, paired=TRUE)

```


The null hypothesis H0 (no difference in mean) is rejected with a mean decrease of 0.63 after 8 weeks.


The permutation test can be applied since we have paired samples


```{r, fig.height=3, fig.width=6}

X = chol_frame$Before
Y = chol_frame$After8weeks
B = 10000
Tstar = numeric(B)
for(i in 1:B){
  xystar = t(apply(cbind(X, Y), 1, sample))
  Tstar[i] = mean(xystar[,1] - xystar[,2])
}
t = mean(X-Y)
pl=sum(Tstar<t)/B;pr=sum(Tstar>t)/B
p = 2*min(pl, pr)
print(p)

```


p is much smaller than 0.95 so there is a significant difference after 8 weeks


# Mann-Whitney can be applied but is less strong - designed for samples from different populations

```{r, fig.height=3, fig.width=6}

wilcox.test(X, Y)

```
again the null hypothesis is rejected.


**c)** 

#we have a discrete sample whose true std is unknown -> student dist


```{r, fig.height=3, fig.width=6}
X <- chol_frame$After8weeks
print("97 percent confidence interval assuming normality:")
mean(X) + qt(c(0.015, 1-0.015), df=length(X)-1) * sd(X) / sqrt(mean(X))

B = 1000
Tstar = numeric(B)
for(i in 1:B){
  Xstar = sample(X, replace=TRUE)
  Tstar[i] = mean(Xstar)
}
Tstar_q = quantile(-Tstar, c(0.015, 1-0.015))
mean(Tstar)
mean(X)
2*mean(X) + Tstar_q

```
with bootstrapping we get a significantly smaller confidence interval.



**d)** 


```{r, fig.height=3, fig.width=6}
B = 1000
Tstar = numeric(B)
theta = seq(from=3, to=12, by=0.001)
num_test_pts = length(theta)
p = numeric(num_test_pts)
for(n in 1:num_test_pts){
  for(i in 1:B){
    Xstar = runif(length(X), 3,theta[n])
    Tstar[i] = max(Xstar)
  }
  t = max(X)
  pl=sum(Tstar<t)/B;pr=sum(Tstar>t)/B
  p[n] = 2*min(pl, pr)
}
plot(theta, p)
print(theta[p>0.95])

```
for theta between 7.83 and 7.88 we have p>0.95 i.e. do not reject H0. 


Kolmogorov Smirnov can be applied by collecting independent samples from U(3, theta)
Since we are now looking at the entire distribution we always reject the null hypothesis

```{r, fig.height=3, fig.width=6}
suppressWarnings({
B = 1000
Tstar = numeric(B)
theta = seq(from=3, to=12, by=0.01)
num_test_pts = length(theta)
p = numeric(num_test_pts)
for(n in 1:num_test_pts){
  
  Y = runif(B, 3, theta)
  k_test = ks.test(X, Y)
  p[n] = k_test$p.value
  
}
print(theta[p>0.95])
})
```



**e)** 
use bootstrapping to get a confidence interval on the median of after8weeks

```{r, fig.height=3, fig.width=6}
X = chol_frame$After8weeks
B = 1000
Tstar = numeric(B)
for(i in 1:B){
  Xstar = sample(X, replace=TRUE)
  Tstar[i] = median(Xstar)
}
Tstar_q = quantile(-Tstar, c(0.05, 1-0.05))
median(X)
2*median(X) + Tstar_q

```


The median has a higher-than 5% chance of being larger than 6 - not statistically significant

again use bootstrapping to get a confidence interval on the 25th percentile

```{r, fig.height=3, fig.width=6}

X = chol_frame$After8weeks
B = 100000
Tstar = numeric(B)
for(i in 1:B){
  Xstar = sample(X, replace=TRUE)
  Tstar[i] = quantile(Xstar, 0.25)
}
Tstar_q = quantile(-Tstar, c(0.05, 1-0.05))
quantile(X, 0.25)
2*quantile(X, 0.25) + Tstar_q
```
















